TimeVaryingDetection_mrmr

### 2024-11-21 code/R/54188_ClassicCustomComparison_mrmr.Rmd
include fit_model function sources

# Mrmr Model Comparison

@@

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting.The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference. The 'changes to mrmr' outlined in this report are primarily changes to the multi-state model, via modifications to the Stan code. The Stan code is not included in this report. Instead, changes are described using the English language. Snippets of essential Stan code are provided, when necessary, in *italics*, following the format (*classic model code*) —\> (*custom model code*).

@@

4.  phi_est - survival probability, estimated between primary periods for the whole population. phi_est[1] is the collective estimated survival probability between the first and second primary periods, phi_est[2] between the second and third, and so on.\*

    -   \*I need to confirm my suspicion that the model generates "boundary estimates" for phi (an estimate before the first primary period, and after the last primary period). I am currently operating under this assumption. If I can't confirm this by myself, maybe I can ask Max.

@@

This section describes the analysis performed to compare the classic and custom models. Both models were fit to the same capture-mark-recapture (CMR) data for the Sierra Nevada yellow-legged frog population at site 54188. Key comparisons include estimates of population abundance (N), recruitment (B), detection probability (p), and survival (phi). To evaluate model performance, I used the **leave-one-out cross-validation (LOO)** method, which assesses model fit while penalizing complexity.

@@

```{r}
# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

@@

```{r}
# Fit the model using mrmr's built in function
classic_mod_54188 <- fit_model_classic_loglik(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

>>>>

# mrmr Model Comparison

@@

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting.The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference. The 'changes to mrmr' outlined in this report are primarily changes to the multi-state model, via modifications to the Stan code. The Stan file is not included in this report. Instead, changes are described using the English language. Snippets of essential Stan code are provided, when necessary, in *italics*, following the format (*classic model Stan code*) —\> (*custom model Stan code*).

@@

4.  phi_est - survival probability, estimated between primary periods for the whole population. phi_est[1] is the collective estimated survival probability between the first and second primary periods, phi_est[2] between the second and third, etc.\*

    -   \*I need to confirm my assumption that the model generates "boundary estimates" for phi (an estimate before the first primary period, and after the last primary period). If I can't confirm this by myself, maybe I can ask Max.

@@

This section describes the analysis performed to compare the classic and custom models. Both models were fit to the same CMR data for the Sierra Nevada yellow-legged frog population at site 54188. Key comparisons include estimates of population abundance (N), recruitment (B), detection probability (p), and survival (phi). To evaluate model performance, I used the leave-one-out cross-validation (LOO) method, which assesses model fit while penalizing complexity.

@@

The fit_model_custom( ) function is identical to mrmr's built in fit_model( ) function, except it draws on the custom .stan file on my local machine, rather than the standard mrmr .stan file. The custom .stan file includes the previously described changes to the parameters, transformed parameters, and generated quantities block.

```{r}
# load the modified fit_model function
source("custom_fit_model_function.R")

# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

@@

The fit_model_classic( ) function is identical to mrmr's built in fit_model( ) function, except it draws on a modified .stan file on my local machine, rather than the standard mrmr .stan file. The modified .stan file includes the previously described changes to the generated quantities block.

```{r}
source("classic_fit_model_function.R")

# Fit the model using mrmr's built in function
classic_mod_54188 <- fit_model_classic(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)
